---
title: "janeeyre"
author: "Katie Larson"
date: "May 22, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

### A clever title

Two years ago Julia Silge posted [an analysis](https://juliasilge.com/blog/you-must-allow-me/) combining natural language processing and Jane Austen. I was thrilled with the idea of analyzing my own favorite pieces of literature, but time got away from me (grad school does that). So here we are, two years later, following Julia's steps to analyze...*Jane Eyre!*

Jane Eyre is in the `novels` dataset found in the `stylo` package.

```{r, include=FALSE}
library(stylo)
data(novels)
janeeyre_text <- novels$CBronte_Jane
remove(novels)
```

The text is divded into paragaphs. To make the rest of the analysis easier, I identified the chapter for each paragraph and removed the title/chapter headings. Jane Eyre has 38 chapters total.

```{r}
# where are the chapters?
chapters_index <- which(startsWith(janeeyre_text, "CHAPTER"))

jane <- data.frame("chapter" = rep(NA, 4021), "text" = rep(NA, 4021))
row <- 0
current_chapter <- 0

# remove title and chapter headings
for(i in 3:length(janeeyre_text)){
  if(i %in% chapters_index){
    current_chapter <- current_chapter + 1
  } else {
    row <- row + 1
    jane[row, 1] <- current_chapter
    jane[row, 2] <- janeeyre_text[i]
  }
}
remove(janeeyre_text)
```

To analyze the emotions in Jane Eyre I took advantage of the [NRC sentiment dictionary](http://saifmohammad.com/WebPages/NRC-Emotion-Lexicon.htm) which is implemented in the `syuzhet` package. The author of `syuzhet` has [a blog](http://www.matthewjockers.net/2015/02/02/syuzhet/) which is very helpful for understanding what the package is capabale of.

The `get_nrc_sentiment` function uses the NRC sentiment dictionary to count occurances of 8 different emotions and two sentiments (positive/negative) within the text. Here is an example using the opening line of the book.

```{r}
library(syuzhet)
get_nrc_sentiment("There was no possibility of taking a walk that day.")
```

Hmm, not a lot of emotion. Let's try another line from the the middle of the book.

```{r}
get_nrc_sentiment("Do you think, because I am poor, obscure, plain, and little, I am soulless and heartless? You think wrong!")
```

There we go! Poor Jane thinks Mr. Rochester is getting married and she's getting shipped off to Ireland because nobody wants her.

Julia makes a great point that these modern tools might not be appropriate for analyzing a text that is over 100 years old. Jane Eyre may also prove challenging because the main characters flirt like this:

>"Am I hideous, Jane?""

>"Very, sir: you always were, you know.""

IM YOU WE WERE US THEY

Nevertheless, Jane is one of the most emotional characters I've ever encountered and I'm excited to see what we find! Now let's get the sentiment scores for each paragrph.

```{r}
jane_nrc <- cbind("chapter" = jane[,1], "paragraph" = 1:nrow(jane), get_nrc_sentiment(jane[,2]))
jane_nrc$negative <- -jane_nrc$negative
```

Now we reshape the data and plot positive/negative scores for each paragraph.

```{r}
library(dplyr)
library(reshape2)
posneg <- jane_nrc %>% select(paragraph, chapter, positive, negative) %>% melt(id = c("paragraph", "chapter"))
names(posneg) <- c("paragraph", "chapter", "sentiment", "value")


annotatetext <- data.frame(x = c(651, 1209, 2245, 2569, 3584, 3768), 
                           y = c(20.3, 23.3, 20.3, 24.3, 20.3, 18.3), 
                           label = c("Helen dies", "Bedroom fire", 
                                     "Kiss", "Wedding", 
                                     "St. John", "Reunited"))
annotatearrow <- data.frame(x = c(651, 1209, 2245, 2569, 3584, 3768), 
                            y1 = c(19,22,19,23,19,17), 
                            y2 = c(12.1, 14.3, 11.4, 14.1, 11.2, 10.5))

ggplot(data = posneg, aes(x=paragraph, y=value, color=sentiment, fill=sentiment)) +
  geom_bar(stat = "identity", position = "dodge") + theme_minimal() +
  ylab("Sentiment") +
  ggtitle(expression(paste("Positive and Negative Sentiment in ", 
                           italic("Jane Eyre")))) +
  theme(legend.title=element_blank()) +
  theme(axis.title.x=element_blank()) +
  theme(axis.ticks.x=element_blank()) +
  theme(axis.text.x=element_blank()) +
  theme(legend.justification=c(1,1), legend.position=c(1, 0.97)) +
  geom_text(data = annotatetext, aes(x,y,label=label), hjust = 0.5, 
            size = 3, inherit.aes = FALSE) +
  geom_segment(data = annotatearrow, aes(x = x, y = y1, xend = x, yend = y2),
               arrow = arrow(length = unit(0.05, "npc")), inherit.aes = FALSE) +
  scale_fill_manual(values = c("goldenrod2", "dodgerblue4")) +
  scale_color_manual(values = c("goldenrod2", "dodgerblue4"))
```

*This is so cool.* The first thing I noticed was the heavy negatives after "wedding" when Jane's Wedding is broken off (because her fiance literally has a crazy wife in the attic) and she runs away, gets lost, and almost starves to death.

It's a good book, I promise.

How does our understanding of the plot change when we look at overall (happy + sad) sentiment? 

```{r}
posneg$overall <- jane_nrc$positive + jane_nrc$negative
ggplot(data = posneg, 
       aes(x = paragraph, y = overall)) +
  geom_bar(stat = "identity", position = "dodge", color = "dodgerblue4") + 
  theme_minimal() +
  ylab("Sentiment") +
  ggtitle(expression(paste("Overall Sentiment in ", italic("Jane Eyre")))) +
  theme(axis.title.x=element_blank()) +
  theme(axis.ticks.x=element_blank()) +
  theme(axis.text.x=element_blank()) +
  theme(legend.justification=c(1,1), legend.position=c(1, 0.71)) +
  geom_text(data = annotatetext, aes(x,y,label=label), hjust = 0.5, 
            size = 3, inherit.aes = FALSE) +
  geom_segment(data = annotatearrow, aes(x = x, y = y1, xend = x, yend = y2),
               arrow = arrow(length = unit(0.05, "npc")), inherit.aes = FALSE)

```

The same negative spikes appear after "Wedding", but I love the additional details that pop up. Jane is mostly happy during her time at Thornfield (between "Helen dies" and "Wedding") but the arrival of the rude, haughty Miss Ingram causes her some distress (between "Bedroom fire" and "Kiss"). The positive spikes before "St. John" show how much Jane loves her new independent life after leaving Thornfield.

Still kinda messy to read. Fouriesr transform. Explain.

```{r}
jane_ft <- as.numeric(get_transformed_values(jane_sentiment$sentiment, 
                                             low_pass_size = 6,
                                             scale_vals = TRUE,
                                             scale_range = FALSE))
jane_ft <- data.frame(cbind(linenumber = 1:100, ft = jane_ft))
ggplot(data = jane_ft, aes(x = linenumber, y = ft)) +
  geom_bar(stat = "identity", alpha = 0.8, color = "dodgerblue4", fill = "dodgerblue4") +
  theme_minimal() +
  ylab("Transformed Sentiment Value") +
  ggtitle(expression(paste("Overall Sentiment in ", italic("Jane Eyre")))) +
  theme(axis.title.x=element_blank()) +
  theme(axis.ticks.x=element_blank()) +
  theme(axis.text.x=element_blank()) +
# helen dies - 651 = 0.1619
# mr rochester room on fire - 1209 = 0.3006715
# kiss under the tree - 2245 = 0.558
# wedding is stopped - 2569 = 0.63889
# st john proposes - 3584 = 0.891
# she proposes to edward, reunited - 3768 = 0.93708
  annotate("text", size = 3, 
           x = c(16, 30, 56, 64, 89, 94), 
           y = c(-0.4, -0.4, -0.55, 1.0, 1.1, 0.6), 
           label = c("Helen", "Fire", "Kiss",
                     "Wedding", "St. John", 
                     "Reunited")) +
  annotate("segment", arrow = arrow(length = unit(0.03, "npc")),
           x = c(16, 30, 56, 64, 89, 94), xend = c(16, 30, 56, 64, 89, 94),
           y = c(-0.36, -0.36, -0.51, 0.92, 1.02, 0.52), 
           yend = c(-0.02, -0.02, -0.02 , 0.02, 0.02, 0.02))
```

Now we're talking! Jane's horrible childhood, the happy engagement, the complete despair when the wedding is broken off, and....sad ending?

Just for fun (Are you sick of Jane Eyre yet? *Never!*) let's look at emotions by chapter.

```{r}
# into chapters
jane_chapters <- data.frame("chapter"=1:38,
                            "text"= rep("NA", 38))
jane_chapters$text <- as.character(jane_chapters$text)
for(i in 1:38){
  index <- which(jane$chapter == i)
  jane_chapters[i,2] <- paste(jane[index,2], sep = '', collapse = '')
}
  

jane_chap_nrc <- cbind("chapter" = jane_chapters[,1], get_nrc_sentiment(jane_chapters[,2]))


emotions <- jane_chap_nrc %>% select(chapter, anger, anticipation, 
                                disgust, fear, joy, sadness, surprise, 
                                trust) %>% melt(id = c("chapter"))
names(emotions) <- c("chapter", "sentiment", "value")
levels(emotions$sentiment) <- c("Anger", "Anticipation", "Disgust", "Fear", 
                                "Joy", "Sadness", "Surprise", "Trust")

emotions$sentiment = factor(emotions$sentiment,levels(emotions$sentiment)[c(5,8,2,7,6,3,4,1)])

library(viridis)
ggplot(data = emotions, aes(x = chapter, 
                            y = sentiment, 
                            fill = value)) +
  geom_tile(color="white", size=0.1) +
  scale_fill_viridis(name="Sentiment\nScore") +
  coord_equal() + 
  labs(x=NULL, y=NULL, 
       title=expression(paste("Sentiment in ", italic("Pride and Prejudice")))) +
  theme(axis.ticks=element_blank(), axis.text.x=element_blank()) +
  scale_x_discrete(expand=c(0,0)) +
  theme(axis.text=element_text(size=6)) +
  theme(panel.border=element_blank()) +
  theme(legend.title=element_text(size=6)) + 
  theme(legend.title.align=1) + 
  theme(legend.text=element_text(size=6)) + 
  theme(legend.position="bottom") + 
  theme(legend.key.size=unit(0.2, "cm")) + 
  theme(legend.key.width=unit(1, "cm"))
```

Bright spot: fear in chapter 27. This is were Jane loses the love of her life, runs away from the only home she's ever known, and nearly starved to death.

Sounds about right.

Somethign to take into considerations: chapters are not the same length!

```{r}
# count number of words
words <- rep(NA, 38)
for(i in 1:38){
  ss <- gsub("[[:punct:]]", "", jane_chapters[i,2])
  words[i] <- length(gregexpr(" ", ss)[[1]])
}


jane_chap_nrc2 <- jane_chap_nrc
for(i in 1:nrow(jane_chap_nrc2)){
  jane_chap_nrc2[i,2:9] <- jane_chap_nrc2[i,2:9]/words[i]
}
emotions2 <- jane_chap_nrc2 %>% select(chapter, anger, anticipation, 
                                     disgust, fear, joy, sadness, surprise, 
                                     trust) %>% melt(id = c("chapter"))
names(emotions2) <- c("chapter", "sentiment", "value")
levels(emotions2$sentiment) <- c("Anger", "Anticipation", "Disgust", "Fear", 
                                "Joy", "Sadness", "Surprise", "Trust")

emotions2$sentiment = factor(emotions2$sentiment,levels(emotions2$sentiment)[c(5,8,2,7,6,3,4,1)])

library(viridis)
ggplot(data = emotions2, aes(x = chapter, 
                            y = sentiment, 
                            fill = value)) +
  geom_tile(color="white", size=0.1) +
  scale_fill_viridis(name="Normalized Score") +
  coord_equal() + 
  labs(x=NULL, y=NULL, 
       title=expression(paste("Emotions in ", italic("Pride and Prejudice")))) +
  theme(axis.ticks=element_blank(), axis.text.x=element_blank()) +
  scale_x_discrete(expand=c(0,0)) +
  theme(axis.text=element_text(size=6)) +
  theme(panel.border=element_blank()) +
  theme(legend.title=element_text(size=6)) + 
  theme(legend.title.align=1) + 
  theme(legend.text=element_text(size=6)) + 
  theme(legend.position="bottom") + 
  theme(legend.key.size=unit(0.2, "cm")) + 
  theme(legend.key.width=unit(1, "cm"))
```


So, something cool: chaper 27 that had the huge "fear" spot has the most words of all the chapters at 11257 - the next largest is 9262.

negative emotion - chapter two, jane is locked in a haunted room as a child.

The happy ending has only 1817, the lowest of any chapter, so, it makes sense that the happy ending gets drowned out?

32 and 31 bright happy spot because, as we saw in fourier transform, jane has a home of her own and is fairly independent. safe. happy.

And finally, THE HAPPY ENDING. YEE HAW.